\begin{sreapi}
\hypertarget{func:esl_stopwatch_Create()}
{\item[ESL\_STOPWATCH * esl\_stopwatch\_Create(void)]}

Creates a new stopwatch.

Returns ptr to a new \ccode{ESL\_STOPWATCH} object; caller is
responsible for free'ing it with 
\ccode{esl\_stopwatch\_Destroy()}.

Throws NULL on allocation failure.


\hypertarget{func:esl_stopwatch_Destroy()}
{\item[void esl\_stopwatch\_Destroy(ESL\_STOPWATCH *w)]}

Frees an \ccode{ESL\_STOPWATCH}.


\hypertarget{func:esl_stopwatch_Start()}
{\item[int esl\_stopwatch\_Start(ESL\_STOPWATCH *w)]}

Start a stopwatch. This sets the base 
for elapsed, cpu, and system time difference
calculations by subsequent calls to
\ccode{esl\_stopwatch\_Stop()}.

Returns \ccode{eslOK} on success.


\hypertarget{func:esl_stopwatch_Stop()}
{\item[int esl\_stopwatch\_Stop(ESL\_STOPWATCH *w)]}

Stop a stopwatch. Record and store elapsed,
cpu, and system time difference relative to the
last call to \ccode{esl\_stopwatch\_Start()}.

Returns \ccode{eslOK} on success.


\hypertarget{func:esl_stopwatch_Display()}
{\item[int esl\_stopwatch\_Display(FILE *fp, ESL\_STOPWATCH *w, char *prefix)]}

Output a usage summary line from a stopped
stopwatch, showing elapsed, cpu, and system time
between the last calls to 
\ccode{esl\_stopwatch\_Start()} and \ccode{esl\_stopwatch\_Stop()}.

The string \ccode{prefix} will be prepended to the output
line. Use \ccode{""} to prepend nothing. If \ccode{prefix} is NULL,
a default \ccode{"CPU Time: "} prefix is used.

For \ccode{prefix} = \ccode{"CPU Time: "} an example output line is:\\
\ccode{CPU Time: 142.55u 7.17s 00:02:29.72 Elapsed: 00:02:35}

Returns \ccode{eslOK} on success.

Throws \ccode{eslEWRITE} on any system write error, such as filled disk.


\hypertarget{func:esl_stopwatch_GetElapsed()}
{\item[double esl\_stopwatch\_GetElapsed(ESL\_STOPWATCH *w)]}

After watch \ccode{w} is Stop()'ed, calling
\ccode{esl\_stopwatch\_GetElapsed(w)} returns the elapsed time
in seconds.

The resolution is system-dependent. 


\hypertarget{func:esl_stopwatch_Include()}
{\item[int esl\_stopwatch\_Include(ESL\_STOPWATCH *master, ESL\_STOPWATCH *w)]}

Merge the cpu and system times from a slave into
a master stopwatch. Both watches must be
stopped, and should not be stopped again unless
You Know What You're Doing.

Elapsed time is not merged. Master is assumed
to be keeping track of the wall clock (real) time,
and the slave/worker watch is ignored.

Useful in at least two cases. One is in 
PVM, where we merge in the stopwatch(es) from separate
process(es) in a cluster. A second is in 
threads, for broken pthreads/times() implementations
that lose track of cpu times used by spawned
threads.

Returns \ccode{eslOK} on success.


\end{sreapi}

