\begin{sreapi}
\hypertarget{func:esl_min_ConjugateGradientDescent()}
{\item[int esl\_min\_ConjugateGradientDescent(ESL\_MIN\_CFG *cfg, double *x, int n, 
       				 double (*func)(double *, int, void *),
				 void (*dfunc)(double *, int, void *, double *),
				 void *prm, double *opt\_fx, ESL\_MIN\_DAT *dat)]}

n-dimensional minimization by conjugate gradient descent.

An initial point is provided by \ccode{x}, a vector of \ccode{n}
components. The caller also provides a function \ccode{*func()} that 
compute the objective function f(x) when called as 
\ccode{(*func)(x, n, prm)}, and a function \ccode{*dfunc()} that can
compute the gradient \ccode{dx} at \ccode{x} when called as 
\ccode{(*dfunc)(x, n, prm, dx)}, given an allocated vector \ccode{dx}
to put the derivative in. Any additional data or fixed
parameters that these functions require are passed by
the void pointer \ccode{prm}.

The first step of each iteration is to try to bracket
the minimum along the current direction. The initial step
size is controlled by \ccode{u[]}; the first step will not exceed 
\ccode{u[i]} for any dimension \ccode{i}. (You can think of \ccode{u} as
being the natural "units" to use along a graph axis, if
you were plotting the objective function.)

The caller also provides an allocated workspace sufficient to
hold four allocated n-vectors. (4 * sizeof(double) * n).

Iterations continue until the objective function has changed
by less than a fraction \ccode{tol}. This should not be set to less than
sqrt(\ccode{DBL\_EPSILON}). 

Upon return, \ccode{x} is the minimum, and \ccode{opt\_fx} (if
provided) is f(x), the function value at \ccode{x}.

Returns \ccode{eslOK} on success. \ccode{x} is updated to be the minimum. \ccode{*opt\_fx},
if provided, contains f(x) at that minimum. \ccode{*opt\_dat}, if
an allocated \ccode{ESL\_MIN\_DAT} was provided, contains a table of 
statistics on the run.

\ccode{eslENOHALT} if it fails to converge in max iterations,
but the final \ccode{x}, \ccode{*opt\_fx}, and \ccode{*opt\_dat} are still provided;
maybe they're good enough, caller can decide.

Throws \ccode{eslERANGE} if the minimum is not finite, which may
indicate a problem in the implementation or choice of \ccode{*func()}.

\ccode{eslEMEM} on allocation failure.
On thrown exceptions, \ccode{*opt\_fx} is eslINFINITY, and \ccode{x} is undefined.



\hypertarget{func:esl_min_dat_Create()}
{\item[ESL\_MIN\_DAT * esl\_min\_dat\_Create(ESL\_MIN\_CFG *cfg)]}

Collects statistics on a CG minimizer run for each
iteration 1..niter, plus the initial position 0.



\end{sreapi}

