\begin{sreapi}
\hypertarget{func:esl_gam_pdf()}
{\item[double esl\_gam\_pdf(double x, double mu, double lambda, double tau)]}

Calculates the gamma PDF $P(X=x)$ given value \ccode{x},
location parameter \ccode{mu}, scale parameter \ccode{lambda}, and shape
parameter \ccode{tau}.


\hypertarget{func:esl_gam_logpdf()}
{\item[double esl\_gam\_logpdf(double x, double mu, double lambda, double tau)]}

Calculates log of the probability density function
for the gamma, $\log P(X=x)$, given value \ccode{x},
location parameter \ccode{mu}, scale parameter \ccode{lambda}, and 
shape parameter \ccode{tau}.


\hypertarget{func:esl_gam_cdf()}
{\item[double esl\_gam\_cdf(double x, double mu, double lambda, double tau)]}

Calculates the cumulative distribution function
for the gamma, $P(X \leq x)$, given value \ccode{x}, 
location parameter \ccode{mu}, scale parameter \ccode{lambda}, and 
shape parameter \ccode{tau}.

(For $\mu=0$, $\lambda = 1$, this is the
incomplete Gamma function $P(\tau,x)$.)


\hypertarget{func:esl_gam_logcdf()}
{\item[double esl\_gam\_logcdf(double x, double mu, double lambda, double tau)]}

Calculates the log of the cumulative distribution function 
for the gamma, $\log P(X \leq x)$, given value \ccode{x}, location
parameter \ccode{mu}, scale parameter \ccode{lambda}, and shape 
parameter \ccode{tau}.


\hypertarget{func:esl_gam_surv()}
{\item[double esl\_gam\_surv(double x, double mu, double lambda, double tau)]}

Calculates the survival function for the gamma, $P(X > x)$,
given value \ccode{x}, location parameter \ccode{mu}, scale parameter 
\ccode{lambda}, and shape parameter \ccode{tau}.


\hypertarget{func:esl_gam_logsurv()}
{\item[double esl\_gam\_logsurv(double x, double mu, double lambda, double tau)]}

Calculates the log of the survival function for the gamma, 
$\log P(X > x)$, given value \ccode{x}, location parameter \ccode{mu},
scale parameter \ccode{lambda}, and shape parameter \ccode{tau}.

Relies on \ccode{esl\_stats\_IncompleteGamma()}, which has limited
dynamic range. Any result of \ccode{ -700 or so will be -infinity.
To fix this, we need a log version of <esl\_stats\_IncompleteGamma()}.


\hypertarget{func:esl_gam_invcdf()}
{\item[double esl\_gam\_invcdf(double p, double mu, double lambda, double tau)]}

Calculates the inverse CDF for a gamma with location
parameter \ccode{mu}, scale parameter \ccode{lambda} and shape
parameter \ccode{tau}, returning the value \ccode{x} at which the
CDF is \ccode{p}.

This inverse CDF is solved by a computationally expensive,
brute force bisection search on the CDF of \ccode{x}.


\hypertarget{func:esl_gam_generic_pdf()}
{\item[double esl\_gam\_generic\_pdf(double x, void *params)]}

Generic-API wrapper around \ccode{esl\_gam\_pdf()}, taking 
a void ptr to a double array containing $\mu$, $\lambda$,
$\tau$ parameters.


\hypertarget{func:esl_gam_generic_cdf()}
{\item[double esl\_gam\_generic\_cdf(double x, void *params)]}

Generic-API wrapper around \ccode{esl\_gam\_cdf()}, taking 
a void ptr to a double array containing $\mu$, $\lambda$,
$\tau$ parameters.


\hypertarget{func:esl_gam_generic_surv()}
{\item[double esl\_gam\_generic\_surv(double x, void *params)]}

Generic-API wrapper around \ccode{esl\_gam\_surv()}, taking 
a void ptr to a double array containing $\mu$, $\lambda$,
$\tau$ parameters.


\hypertarget{func:esl_gam_generic_invcdf()}
{\item[double esl\_gam\_generic\_invcdf(double x, void *params)]}

Generic-API wrapper around \ccode{esl\_gam\_invcdf()}, taking 
a void ptr to a double array containing $\mu$, $\lambda$,
$\tau$ parameters.


\hypertarget{func:esl_gam_Plot()}
{\item[int esl\_gam\_Plot(FILE *fp, double mu, double lambda, double tau,
	     double (*func)(double x, double mu, double lambda, double tau), 
	     double xmin, double xmax, double xstep)]}

Plot some gamma distribution function \ccode{func} (for instance,
\ccode{esl\_gam\_pdf()}) for parameters \ccode{mu}, \ccode{lambda}, and \ccode{tau}, for
a range of values x from \ccode{xmin} to \ccode{xmax} in steps of \ccode{xstep};
output to an open stream \ccode{fp} in xmgrace XY input format.

Returns \ccode{eslOK} on success.

Throws \ccode{eslEWRITE} on any system write error, such as a filled disk.


\hypertarget{func:esl_gam_Sample()}
{\item[double esl\_gam\_Sample(ESL\_RANDOMNESS *r, double mu, double lambda, double tau)]}

Sample a gamma-distributed random variate.


\hypertarget{func:esl_gam_FitComplete()}
{\item[int esl\_gam\_FitComplete(double *x, int n, double mu, double *ret\_lambda, double *ret\_tau)]}

Given complete data consisting of \ccode{n} samples \ccode{x[0]..x[n-1]},
and a known location parameter \ccode{mu}, determine and return
maximum likelihood parameters \ccode{ret\_lambda} and \ccode{ret\_tau}.

Returns \ccode{eslOK} on success.

Throws \ccode{eslENOHALT} if bracketing or bisection fails;
\ccode{eslEINVAL} if data cannot be gamma distributed (some \ccode{x[i] < mu},
or zero variance in x).



\hypertarget{func:esl_gam_FitCompleteBinned()}
{\item[int esl\_gam\_FitCompleteBinned(ESL\_HISTOGRAM *g, double *ret\_mu, double *ret\_lambda, double *ret\_tau)]}

Fit a complete exponential distribution to the observed
binned data in a histogram \ccode{g}, where each
bin i holds some number of observed samples x with values from 
lower bound l to upper bound u (that is, $l < x \leq u$);
determine and return maximum likelihood estimates for the
parameters $\mu, \lambda, \tau$ and 
return them in \ccode{*ret\_mu}, \ccode{*ret\_lambda}, \ccode{*ret\_tau}.

Unlike the \ccode{esl\_exp\_FitCompleteBinned()} case where the
ML fit optimizes $\sum_i n_i \log P(a_i \leq x < b_i)$
where $a_i \leq b_i$ are the bounds of bin i with
occupancy $n_i$, here we take the approximation that
$c_i = a_i + 0.5*(b_i-a_i)$ and optimize $\log P(a_i
\leq x < b_i) \simeq \log(w) + \log P(x=c_i)$.

Since $b_i-a_i = w$ is fixed, optimizing the above
becomes equivalent to optimizing $\sum_i n_i * log P(x=c_i)$.

The optimization is then equivalent to the non-binned case,
but subsituting in averages such as $\sum_i x(i)$ by
$\sum_i n_i*c_i i$, and so forth.

If the binned data in \ccode{g} were set to focus on 
a tail by virtual censoring, the "complete" exponential is 
fitted to this tail. The caller then also needs to
remember what fraction of the probability mass was in this
tail.

Returns \ccode{eslOK} on success.

Throws \ccode{eslENOHALT} if bracketing or bisection fails;
\ccode{eslEINVAL} if data cannot be gamma distributed (some \ccode{x[i] < mu},
or zero variance in x).




\end{sreapi}

